---
title: "Appendix B"
subtitle: "USE CASE STUDY: AMERICAN SAMOA"
authors:
  - name: Julie Vercelloni
    corresponding: true
    email: j.vercelloni@aims.gov.au
    affiliations:
      - ref: aims
      - ref: qut
  - name: Murray Logan
    affiliations:
      - ref: aims
  - name: Andrew Zammit‐Mangion
    affiliations:
      - ref: UOW
      - ref: KAUST
  - name: Matthew Sainsbury‐Dale
    affiliations:
      - ref: UOW
      - ref: KAUST
  - name: Britta Schaffelke
    affiliations:
      - ref: aims
  - name: Kerrie Mengersen
    affiliations:
      - ref: qut
  - name: Manuel González‐Rivero
    affiliations:
      - ref: aims
affiliations:
  - id: aims
    name: Australian Institute of Marine Science, Townsville, Australia
  - id: qut
    name: Centre for Data Science, Queensland University of Technology, Brisbane, Australia
  - id: UOW
    name: School of Mathematics and Applied Statistics, University of Wollongong, Wollongong, Australia
  - id: KAUST
    name: King Abdullah University of Science and Technology, Thuwal, Saudi Arabia
engine: knitr
format:
  html:
    toc: true
    toc-location: left
    title-block-banner: true
    toc-depth: 3
    highlight-style: atom-one
    embed-resources: true
    theme:
      light: flatly
      dark: darkly
    code-overflow: wrap
    code-fold: false
    number-sections: true
    number-depth: 2
    shift-heading-level-by: -1
    crossref:
      lst-title: Code listing
    fig-align: center
    text-align: center
    acm-html: default
css: styles.css
editor:
  markdown:
    wrap: 72
---

```{r setup, chunkOpts, echo = FALSE}
knitr::opts_chunk$set(cache = FALSE, message = FALSE, warning = FALSE)
```
::: {.callout-important}
To be displayed on a big screen. 
:::


```{r}
#| include: false
#| echo: false
#| eval: true
rm(list=ls())
setwd("c:/Users/jvercell/OneDrive - Australian Institute of Marine Science/AIMS/01_Research projects/ReefCloud/SP_models/FRK_dev/git_content/Appendix_B/scripts")

source("../R/functions.R")

library(tidyverse)
library(dplyr)
library(purrr)
library(FRK) # remotes::install_github('andrewzm/FRK', ref = 'sumaggregation')
library(spacetime)
library(lubridate)
library(sf)
library(sp)
library(patchwork)
library(ggpubr)
library(tidybayes)
library(ggridges)
library(ggspatial)
library(ggrepel)
library(glue)
library(kableExtra)
library(wesanderson)
library(rnaturalearth)

appendix_dir <- "../figures" 
```{r}
#| include: false
#| echo: false
#| eval: false

###########################################
############################ RUN FRK MODELS (!! it is taking hours !!)
###########################################

## Load tables
load("../data/covariates_full_tier5.RData")
load("../data/tiers.lookup.RData")
load("../data/tier5.sf.RData")

## Define focal tier
BY_TIER = 5 

## Import reef layer 
load("../data/reef_layer.sf.RData")

# central GBR only
depth_strata <- c("shallow", "deep")  

# Define focal tier
FOCAL_TIER <- paste0('Tier', as.numeric(BY_TIER) - 1)
data.grp.agg <- read.csv("../data/reef_data_aggregated.csv") 

# Plots raw data 
g <- ggplot(
  data.grp.agg |> 
    mutate(Tier5 = as.character(Tier5)),
  aes(x = fYEAR, y = COUNT / TOTAL, group = (Tier5))
) +
  geom_line(color = "grey40", alpha = 0.6, linewidth = 0.6) +
  geom_point(aes(color = Tier5), size = 1.8, alpha = 0.8) +
  theme_pubr() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_line(color = "grey85"),
    panel.grid.minor = element_blank()
  ) +
  labs(
    x = "Year",
    y = "Proportion (COUNT / TOTAL)"
  )

TIER <- unique(data.grp.agg$Tier4)


#########################
#### START LOOP BY TIER
#########################

#for (i in seq_along(depth_strata)) {

#STRATA <<- as.character(depth_strata[i])

data.grp.tier <- data.grp.agg |>
     # dplyr::filter(Depth_strata == STRATA) |>
      dplyr::mutate(Tier5 = as.character(Tier5))


tier.sf.joined <- tier.sf |>
    dplyr::left_join(
      tiers.lookup |> dplyr::select(-reef_area, -tier_id),
      by = c("Tier5" = "Tier5")
    ) |> 
      dplyr::filter(!!sym(FOCAL_TIER) == TIER)

## Load covariate layers
full_cov_raw <- full_cov |> 
      dplyr::filter(Tier5 %in% tier.sf.joined$Tier5) |>
      dplyr::rename(fYEAR = year) |>
      dplyr::filter(between(fYEAR, min(data.grp.tier$fYEAR), max(data.grp.tier$fYEAR)))

full_cov_raw_dat <- full_cov_raw %>%
  filter(Tier5 %in% data.grp.tier$Tier5) %>%
  filter(fYEAR %in% unique(data.grp.tier$fYEAR))


## Apply quality control thresholds
out_cycl <- quantile(full_cov_raw$max_cyc, probs = 0.975, na.rm = T)
out_dhw  <- quantile(full_cov_raw$max_dhw, probs = 0.975, na.rm = T)
    
HexPred_sf <- full_cov_raw |>
      dplyr::mutate(As.Data = ifelse(Tier5 %in% data.grp.tier$Tier5, "Yes", "No")) |>
      dplyr::mutate(across(matches("^max_cyc.*"), ~ ifelse(.x >= out_cycl & As.Data == "No", NA, .x))) |>
      dplyr::mutate(across(matches("^max_dhw.*"), ~ ifelse(.x >= out_dhw & As.Data == "No", NA, .x)))

# ## Select covariates using info across regions 
# selected_covar <- select_covariates(HexPred_sf) 
# selected_covar <- selected_covar[!grepl("\\.(lag3|lag5)$", selected_covar)]

# ## Select covariates using info at data locations
# selected_covar <- select_covariates(full_cov_raw_dat)

# ## Check for collinearity 

# selected_covar <- filter_non_collinear(full_cov_raw_dat, selected_covar, threshold = 0.7)

## Make reefid

covs.hexpred_tier_sf_v2_prep <- make_reefid(tier.sf.joined, HexPred_sf, reef_layer.sf) 

HexPred_reefid <- covs.hexpred_tier_sf_v2_prep |>
      dplyr::group_by(Tier5) |>
      dplyr::summarise(reefid = paste0(reefid, collapse = "_")) |>
      ungroup()

HexPred_reefid2 <- inner_join(HexPred_sf |> data.frame(), HexPred_reefid) |>
      dplyr::group_by(Tier5, fYEAR) |>
      dplyr::filter(row_number() == 1) |>
      dplyr::mutate(across(everything(), ~ replace(.x, is.na(.x), 0))) |>
      sf::st_as_sf(sf_column_name = "geometry")

# In case observations are not yet aggregated (FRK takes care of it automatically)
data.grp.tier.ready <- rm_obs_outside(data.grp.tier, HexPred_reefid2)    

## Test if more than one reef in the final data 
test_reefid <- HexPred_reefid2 %>% filter(Tier5 %in% data.grp.tier.ready$Tier5)

## For dev only
#obj_frk <- frk_prep_2L(data.grp.tier.ready, HexPred_reefid2)  

obj_frk <- frk_prep_3L(data.grp.tier.ready, HexPred_reefid2)  

## Define model formula
# model_formula <- if (length(selected_covar) == 0 && length(unique(test_reefid$reefid)) > 1) {
#         as.formula("COUNT ~ 1 + (1 | reefid)")
#        } else if (length(selected_covar) == 0 && length(unique(test_reefid$reefid)) == 1) {
#         as.formula("COUNT ~ 1")
#       } else if (length(selected_covar) != 0 && length(unique(test_reefid$reefid)) == 1) {
#         as.formula( paste("COUNT ~ 1 +", paste(selected_covar, collapse = " + ")))
#       } else {
#        as.formula(paste("COUNT ~ 1 + (1 | reefid) +", paste(selected_covar, collapse = " + ")))
#   }

model_formula <- COUNT ~ 1 + (1 | reefid) + max_cyc + max_dhw 

## Fit FRK model
M <- FRK(
      f = model_formula,
      data = list(obj_frk$STObj),
      BAUs = obj_frk$ST_BAUs,
      basis = obj_frk$basis,
      response = "binomial",
      link = "logit",
      K_type = "precision",
      method = "TMB",
      est_error = FALSE
    )

##############################
#### Predict & summarise
##############################
pred <- FRK::predict(M, type = "mean", nsim = 1000)

post_dist_df <- as.data.frame(pred$MC$mu_samples) |>
      dplyr::mutate(fYEAR = obj_frk$ST_BAUs@data$fYEAR,
                    Tier5 = obj_frk$ST_BAUs@data$Tier5,
                    id_loc = row_number()) |>
      tidyr::pivot_longer(!c(fYEAR, Tier5, id_loc), names_to = "draw", values_to = "pred") |>
      dplyr::mutate(model_name = "FRK")

tier.sf.joined$Tier5 <- as.factor(tier.sf.joined$Tier5)

pred_sum_sf <- post_dist_df |> group_by(fYEAR, Tier5) |>
      ggdist::median_hdci(pred) |>
      dplyr::inner_join(tier.sf.joined |> dplyr::select(geometry, Tier5)) |>
      sf::st_as_sf(sf_column_name = "geometry") |>
      dplyr::mutate(Unc = .upper - .lower,
                    Tier5_fYEAR = paste0(Tier5, fYEAR))

###############################
##### Save outputs
###############################
saveRDS(
      list(
        form = model_formula,
        pred_sum_sf = pred_sum_sf,
        post_dist_df = post_dist_df,
        data.grp.tier = data.grp.tier,
        M = M,
        obj_frk = obj_frk,
        HexPred_reefid2 = HexPred_reefid2
      ),
      file = paste0("../data/FRK_shallow.RData")
    )
#}
```


```{r}
#| include: false
#| echo: false
#| eval: true
 
# Read model outputs 
model_files <- c(
  "../data/FRK_shallow.RData"
)

mod_ <- map(model_files, readRDS)

names(mod_) <- map2_chr(
  mod_,
  model_files,
  ~ paste0(
      "FRK",
      "_",
      unique(.x$data.grp.tier$Depth_strata)
    )
)
```

```{r}
#| include: false
#| echo: false
#| eval: true

# Vizualise row data 
plots_hov <- imap(
  mod_,
  ~ data_viz(.x$data.grp.tier, .y)
)
```

```{r}
#| include: false
#| echo: false
#| eval: true

# Vizualise disturbances

load("../data/tiers.lookup.RData")

sf_use_s2(FALSE) 
plots_dhw <- imap(
  mod_,
  ~ make_disturbance_plot(.x,
                          var = "max_dhw",
                          dist_label = "DHW",
                          fill_name = "Mean Degree Heating Weeks",
                          palette_name = "lapaz",
                          direction = 1) #+ 
   # ggtitle(sub("_.*", "", .y))
)

plots_cyc <- imap(
  mod_,
  ~ make_disturbance_plot(.x,
                          var = "max_cyc",
                          dist_label = "cyclone exposure",
                          fill_name = "Mean Cyclone Exposure (in hours)",
                          palette_name = "hawaii",
                          direction = -1) #+ 
   # ggtitle(sub("_.*", "", .y))
)

dist_shallow <- plots_dhw[[1]] +  plots_cyc[[1]] + 
   plot_layout(ncol = 2, axis_titles = "collect") +  
   plot_annotation(tag_levels = "a", tag_suffix = ')', 
  title = "American Samoa"
)

```

```{r}
#| include: false
#| echo: false
#| eval: true

# Plot maps 

sf::sf_use_s2(FALSE) 

unique_hex <- mod_ %>%
  imap_dfr(~ {
    df_hex <- .x$HexPred_reefid2 %>%
      group_by(Tier5) %>%
      filter(row_number() == 1) %>%
      dplyr::select(Tier5, geometry) %>%
      ungroup()

    df_hex <- df_hex %>%
      mutate(
        Depth_strata = unique(.x$data.grp.tier$Depth_strata)[1]
      )
    df_hex
  }) %>%
  distinct(Tier5, .keep_all = TRUE)

tier5_list <- imap_dfr(mod_, ~ {
  tibble(
    Tier5  = unique(.x$data.grp.tier$Tier5),
    Region = sub("_.*", "", .y)
  )
})

load("../data/tiers.lookup.RData")
load("../data/tier5.sf.RData")

tier.sf.joined <- tier.sf |>
    dplyr::left_join(
      tiers.lookup |> dplyr::select(-reef_area, -tier_id),
      by = c("Tier5" = "Tier5")
    )

## Whole American Samoa 
roi_bbox  <- st_bbox(tier.sf.joined)
roi_poly  <- st_as_sfc(roi_bbox)
asm <- ne_countries(scale = "medium", country = "American Samoa", returnclass = "sf")

# Crop coastline to ROI
asm_crop <- st_intersection(asm, roi_poly)

# lat_box <- st_bbox(c(
#   xmin = 142,
#   xmax = 152,
#   ymin = -20.7,
#   ymax = -15.4
# ), crs = st_crs(tier.sf.joined))

# lat_box_sf <- st_as_sfc(lat_box)

# tier_filtered <- st_intersection(tier.sf.joined, lat_box_sf) %>% filter(!Tier4 == "1825")
# tier_filtered %>% group_by(Tier4) %>% count()
# tier4_polygons <- tier.sf.joined %>% filter(Tier4 %in% unique(tier_filtered$Tier4)) %>% group_by(Tier4) %>% count()
# tier4_polygons_filtered <- tier_filtered %>% group_by(Tier4) %>% count()

# tier4_south  <- unique_hex %>%
#   mutate(Tier4_NRM = case_when(Tier4 == "1819" ~ "Wet Tropics",
#                                Tier4 == "1808" ~ "Burdekin",
#                                Tier4 == "1820" ~ "Mackay Whitsunday",
#                                TRUE ~ NA_character_))
# tier4_south$Tier4_NRM <- as.factor(tier4_south$Tier4_NRM)
# tier4_south$Tier4_NRM <- factor(tier4_south$Tier4_NRM, levels = c("Wet Tropics", "Burdekin", "Mackay Whitsunday"))

# # Bounding box of the filtered tiers
# # bbox_filtered <- st_bbox(tier_filtered)
# # bbox_poly <- st_as_sfc(bbox_filtered)

# roi_bbox  <- st_bbox(tier4_polygons)
# roi_poly  <- st_as_sfc(roi_bbox)
# aus <- ne_countries(scale = "medium", country = "Australia", returnclass = "sf")

# # Crop coastline to ROI
# aus_crop <- st_intersection(aus, roi_poly)

tier4_south_showed <- unique_hex %>% 
  filter(Tier5 %in% tier5_list$Tier5) %>%
  st_centroid() %>%                     
  mutate(
    lon = st_coordinates(.)[,1],
    lat = st_coordinates(.)[,2]
  ) %>%
  st_drop_geometry()

add_islands <- tibble(
  NAME = c("Tutuila", "Ofu-Olosega", "Ta‘ū", "Rose Atoll", "Swains Island"),
  lon  = c(-170.700, -169.630, -169.500, -168.200, -171.076),
  lat  = c(-14.275, -14.190, -14.240, -14.550, -11.053)
) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  mutate(
    lon = st_coordinates(geometry)[,1],
    lat = st_coordinates(geometry)[,2]
  ) 
## Plots per islands 

add_islands <- add_islands %>%
  mutate(buffer_deg = c(0.2, 0.08, 0.0879, 0.08, 0.05)) # one buffer per island


# Generate plots
plots_per_island <- plot_islands_custom_buffer(add_islands)$plots

# Example: display a specific island

p1 <- plots_per_island[["Tutuila"]]
p2 <- plots_per_island[["Ofu-Olosega"]]
p3 <- plots_per_island[["Ta‘ū"]]
p4 <- plots_per_island[["Rose Atoll"]]
p5 <- plots_per_island[["Swains Island"]]

p_islands <- p1 + p2 + p3 + p4 + p5 +
   plot_layout(nrow = 2, axis_titles = "collect") +  
   plot_annotation(tag_levels = "a", tag_suffix = ')'
)
p_islands

ggsave(p_islands, filename = "../figures/ASM_region_islands.png", width = 12, height = 10)

## All together
hex_island <- plot_islands_custom_buffer(add_islands)$unique_hex_island


data_island <- mod_[[1]]$data.grp.tier  %>%
left_join(hex_island)

p1 <- ggplot() +
  geom_sf(data = asm_crop,
          fill = "grey95",
          colour = "black",
          linewidth = 0.6) +
  geom_sf(data = tier.sf.joined,
          fill = NA,
          colour = "black",
          alpha = 0.2) +
  geom_sf(data = hex_island,
          aes(fill = NAME),
          alpha = 0.8,
          linewidth = 0.05 ) +
  geom_text_repel(data = add_islands,
                   aes(x = lon, y = lat, label = NAME),
                   size = 3,
                   nudge_x = -.3,
                   box.padding = 0.1,
                   nudge_y = -.1) +
  geom_point(data = tier4_south_showed,
           aes(x = lon, y = lat),
           shape = 21, fill = "red", colour = "black", size = 1, alpha = 0.6) +
  coord_sf(xlim = c(roi_bbox["xmin"], roi_bbox["xmax"]),
           ylim = c(roi_bbox["ymin"], roi_bbox["ymax"])) +
  ggpubr::theme_pubr() +
  xlab("Longitude") +
  ylab("Latitude") + 
  annotation_scale(
  location = "bl",        
  width_hint = 0.3
) +
annotation_north_arrow(
  location = "tr",         
  which_north = "true",
  style = north_arrow_fancy_orienteering,
  pad_x = unit(0.1, "cm"),
  pad_y = unit(.02, "cm")
) +
      scale_fill_manual(values = c(
  "Tutuila"       = "#5b2a86",  
  "Ofu-Olosega"   = "#2c7fb8",  
  "Ta‘ū"          = "#41b6c4",  
  "Rose Atoll"    = "#7fcdbb", 
  "Swains Island" = "#f768a1"
)) + theme(legend.position = "none")

ggsave(p1, filename = "../figures/ASM_region_appendix.png", width = 8, height = 8)

## Paper 
bbox_filtered <- st_bbox(unique_hex)
bbox_poly <- st_as_sfc(bbox_filtered) %>%
  st_buffer(0.05) 

tier5_marked <- c("1912", "1902", "1936") #"10746",  "11305", 
tier4_south_showed <- unique_hex %>% 
  filter(Tier5 %in% tier5_marked) %>%
  st_centroid() %>%                     
  mutate(
    lon = st_coordinates(.)[,1],
    lat = st_coordinates(.)[,2]
  ) %>%
  st_drop_geometry()

pp1 <- ggplot() +
  geom_sf(data = asm_crop,
          fill = "grey95",
          colour = "black",
          linewidth = 0.6) +
  geom_sf(data = tier.sf.joined,
          fill = NA,
          colour = "black",
          alpha = 0.2) +
  geom_sf(data = hex_island,
          aes(fill = NAME),
          colour = "#4D4D4D",
          alpha = 0.8,
          linewidth = 0.05 ) +
  geom_text_repel(data = tier4_south_showed,
                  aes(x = lon, y = lat, label = Tier5),
                  size = 4,
                 nudge_x = -.1,
                   box.padding = 0.1,
                   nudge_y = .3) +
  geom_sf(data = bbox_poly,
        aes(colour = "American Samoa"),
          fill = NA,
          linewidth = 0.9) +
  coord_sf(xlim = c(roi_bbox["xmin"], roi_bbox["xmax"]),
           ylim = c(roi_bbox["ymin"], roi_bbox["ymax"])) +
  ggpubr::theme_pubr() +
  xlab("Longitude") +
  ylab("Latitude") + 
  scale_colour_manual(
  name = NULL,
  values = c("American Samoa" = "#9B2226")
  ) + 
  annotation_scale(
  location = "bl",        
  width_hint = 0.3
  ) +
  annotation_north_arrow(
  location = "bl",         
  which_north = "true",
  style = north_arrow_fancy_orienteering,
  pad_x = unit(0.1, "cm"),
  pad_y = unit(.02, "cm")
  ) +
  scale_fill_manual("Island", values = c(
  "Tutuila"       = "#5b2a86",  
  "Ofu-Olosega"   = "#2c7fb8", 
  "Ta‘ū"          = "#41b6c4", 
  "Rose Atoll"    = "#7fcdbb",  
  "Swains Island" = "#f768a1"
)) + theme(legend.position = "top")
pp1 

ggsave(pp1, filename = "../figures/ASM_region.png", width = 10, height = 10)

## Get info region 
```
### American Samoa {.unnumbered}

American Samoa comprises five islands, representing a total reef area of 59.2 km², with 75% of tiers containing data. This case study focuses on islands belonging to the single ecoregion that encompasses the entire country (@fig-map-1) and its islands (@fig-map-2).

Coral cover observations were extracted from the ReefCloud database on 15 March 2026. A total of 507 transects from 507 sites collected between 2015 and 2023 by the NOAA's U.S. Pacific Islands National Coral Reef Monitoring Program (@tbl-1) at depths varying 3 and 10 metres. At the data-tier level, observed coral cover varied from 0.5\% to 69.6\% with a median coral cover values estimated at 25.3\% between 2015 to 2023 (@fig-data-1). 

The extracted disturbance values indicated multiple potential events, including five years with non-zero Degree Heating Weeks (DHW) and two years with detected exposure to cyclonic waves (@fig-dist-1). Disturbance events were infrequently captured in the data tiers: only the 2015 heat-stress event was recorded at higher intensity, and Cyclone Gita (2018) at a comparable intensity to the regional signal.

```{r, fig.width=8, fig.height=10}
#| include: true
#| echo: false
#| eval: true
#| label: fig-map-1
#| fig-cap: Locations of the five islands across the American Samoa. Red dots indicates the tiers containing coral cover observations between 2015-2023. 
p1
```

```{r, fig.width=8, fig.height=6}
#| include: true
#| echo: false
#| eval: true
#| label: fig-map-2
#| fig-cap: Zoom-in on the islands. Red dots indicates the tiers containing coral cover observations between 2015-2023. 
p_islands
```

```{r}
#| label: tbl-1
#| eval: true
#| echo: false
#| tbl-cap: Summary of public datasets available in the American Samoa. Percentages are based on the absolute counts provided in parentheses. Monitoring programs are classified by temporal resolution with short-term (0–5 years), medium-term (6–15 years), and long-term (>15 years).   
print_attr_plots(appendix_dir, "summary_table_ASM")  
```


```{r, fig.width=5, fig.height=6}
#| include: true
#| echo: false
#| eval: true
#| label: fig-data-1
#| fig-cap: Temporal pattern of mean coral cover (%) by data‐tiers in the American Samoa (2015–2023). White gaps represent missing data.
plots_hov[[1]]
```

```{r, fig.width=12, fig.height=14}
#| include: true
#| echo: false
#| eval: true
#| label: fig-dist-1
#| fig-cap: Proportion of regional areas affected by disturbances across two data levels all‐tier and data‐tier for each year between 2015 and 2023 (in the case of data‐tier, the proportion is calculated with respect to the number of all‐tiers). Colours indicate a) average Degree Heating Weeks (DHW) and b) average exposure to cyclone waves. 
dist_shallow
```

```{r}
#| include: false
#| echo: false
#| eval: false

### Spatial weighted aggregation 

################################ NRM trends
load("../data/tiers.lookup.RData")
load("../data/tier5.sf.RData")

# Extract model outputs at the different spatial scales 
res <- spatial_agg_prep()

post_dist_df_tier5 <- res$post_dist_df_tier5
post_dist_df_all   <- res$post_dist_df_all
```


```{r}
#| include: false
#| echo: false
#| eval: false

#### Local scale {.unnumbered}
pred_local_shallow <- spatial_agg(post_dist_df_tier5, post_dist_df_all, tiers.lookup, tier.sf, tier_col = "Tier5")

unique_hex_Tutuila <- plot_islands_custom_buffer(add_islands)$unique_hex_island %>%
  filter(NAME == "Tutuila") 

list_tier5 <-  unique(pred_local_shallow$pred_table$Tier5)

# Select data-tiers

mean_obs <- mod_[[1]]$data.grp.tier   %>%
 group_by(Tier5, fYEAR) %>%
 summarize(COUNT_sum = sum(COUNT),
           TOTAL_sum = sum(TOTAL)) %>%
 ungroup() %>% 
    group_by(fYEAR, Tier5) %>%
    summarize(
      mean_cover = mean(COUNT_sum / TOTAL_sum) * 100,
      .groups = "drop"
    ) %>%
    mutate(
      fYEAR = as.factor(fYEAR),
      Tier5 = as.factor(Tier5)
    ) %>%
  mutate(Year = as.integer(as.character(fYEAR)))

mean_obs_Tutuila <- mean_obs %>%
 filter(Tier5 %in% unique_hex_Tutuila$Tier5) %>%
 group_by(Tier5) %>% count() %>% arrange(desc(n))

plots_with_data <- pred_local_shallow$pred_table %>%
     dplyr::left_join(
        tiers.lookup |> dplyr::select(-reef_area, -tier_id),
        by = c("Tier5" = "Tier5")
       ) %>% 
  droplevels() %>%
  filter(Tier5 %in% unique(mean_obs$Tier5)) %>%
  mutate(Year = as.integer(Year)) %>%
  left_join(mean_obs, by = c("Tier5" = "Tier5", "Year" = "Year")) %>%
  mutate( 
    arrow_symbol = case_when(
      Change == "Up"   ~ "↑",
      Change == "Down" ~ "↓",
      Change == "Flat" ~ "→",
      TRUE ~ ""
    )
  ) %>%
  split(.$Tier5) %>%
  imap(~{
    ggplot(.x) +
      geom_ribbon(aes(x = Year, ymin = Lower*100, ymax = Upper*100),
                  alpha = 0.2,
                  fill = "#5b2a86") +
      geom_line(aes(x = Year, y = Median*100),
                linewidth = 1.1,
                colour = "#4D4D4D") +
      geom_point(aes(x = Year, y = mean_cover), 
                fill = "grey66", 
                col = "black", 
                size = 2.7, 
                alpha = .6, 
                shape = 20) + 
      geom_text(
        aes(x = Year,
            y = Upper*100 + 5,
            label = arrow_symbol),
        size = 6
      ) +
      labs(
        title = .y,
        x = "Year",
        y = "Coral cover (%)"
      ) +
      ylim(0, 80) +
      theme_pubr() +
      theme(
        axis.text  = element_text(size = 13),
        axis.title = element_text(size = 15),
        plot.title = element_text(size = 16),
        legend.position = "none"
      )
  })

p1 <- plots_with_data[[which(list_tier5 == 1912)]]
p2 <- plots_with_data[[which(list_tier5 == 1927)]]
p3 <- plots_with_data[[which(list_tier5 == 1936)]]

p_data <- p1 + p2 + p3 +
   plot_layout(nrow = 2, axis_titles = "collect") +  
   plot_annotation(tag_levels = "a", tag_suffix = ')', 
  title = "Tutuila"
)
ggsave(p1, filename = "../figures/data_tier_example1.png",
        width=6, height=5)

ggsave(p2, filename = "../figures/data_tier_example2.png",
        width=6, height=5)

ggsave(p3, filename = "../figures/data_tier_example3.png",
        width=6, height=5)


# All the tier5

setdiff(unique(unique_hex_Tutuila$Tier5),
        unique(mean_obs_Tutuila$Tier5))

plots <- pred_local_shallow$pred_table %>%
     dplyr::left_join(
        tiers.lookup |> dplyr::select(-reef_area, -tier_id),
        by = c("Tier5" = "Tier5")
       ) %>% 
  droplevels() %>%
  mutate( 
    arrow_symbol = case_when(
      Change == "Up"   ~ "↑",
      Change == "Down" ~ "↓",
      Change == "Flat" ~ "→",
      TRUE ~ ""
    )
  ) %>%
  split(.$Tier5) %>%
  imap(~{
    ggplot(.x) +
      geom_ribbon(aes(x = Year, ymin = Lower*100, ymax = Upper*100),
                  alpha = 0.2,
                  fill = "#5b2a86") +
      geom_line(aes(x = Year, y = Median*100),
                linewidth = 1.1,
                colour = "#4D4D4D") +
      geom_text(
        aes(x = Year,
            y = Upper*100 + 3,
            label = arrow_symbol),
        size = 6
      ) +
      labs(
        title = .y,
        x = "Year",
        y = "Coral cover (%)"
      ) +
    #  ylim(0, 80) +
      theme_pubr() +
      theme(
        axis.text  = element_text(size = 13),
        axis.title = element_text(size = 15),
        plot.title = element_text(size = 16),
        legend.position = "none"
      )
  })

p0 <- plots[[which(list_tier5 == 1902)]]
p0.1 <- plots[[which(list_tier5 == 1929)]]
p0.3 <- plots[[which(list_tier5 == 1950)]]

p_datano <- p0 + p0.1 + p0.3 +
   plot_layout(nrow = 2, axis_titles = "collect") +  
   plot_annotation(tag_levels = "a", tag_suffix = ')', 
  title = "Tutuila"
)

ggsave(p0, filename = "../figures/pred_tier_example1.png",
        width=6, height=5)

ggsave(p0.1, filename = "../figures/pred_tier_example2.png",
        width=6, height=5)

ggsave(p0.3, filename = "../figures/pred_tier_example3.png",
        width=6, height=5)

```

```{r, fig.width=8, fig.height=6}
#| include: true
#| echo: false
#| eval: false
#| label: fig-pred-data
#| fig-cap: Examples of three data-tiers with estimated coral trends from 2015 to 2023. Black lines represent mean predicted coral cover, with shaded areas indicating predictive intervals. Dots show observed coral cover averaged at the tier level for each year, which were used as the response variable in the spatio-temporal model. 
p_data
```
```{r, fig.width=8, fig.height=6}
#| include: true
#| echo: false
#| eval: false
#| label: fig-pred-datano
#| fig-cap: Examples of estimated coral trends at predictive-tiers from 2015 to 2023. Black lines represent mean predicted coral cover, with shaded areas indicating predictive intervals.
p_datano
```

```{r}
#| include: false
#| echo: false
#| eval: false

#### Island scale {.unnumbered}

unique_hex_island <- plot_islands_custom_buffer(add_islands)$unique_hex_island

pred_island <- spatial_agg_island(post_dist_df_all, tiers.lookup, tier.sf, unique_hex_island)

plots <- pred_island$pred_table %>%
  droplevels() %>%
  mutate(
    arrow_symbol = case_when(
      Change == "Up"   ~ "↑",
      Change == "Down" ~ "↓",
      Change == "Flat" ~ "→",
      TRUE ~ ""
    )
  ) %>%
  split(.$NAME) %>%
  imap(~{
    ggplot(.x) +
      geom_ribbon(aes(x = Year, ymin = Lower*100, ymax = Upper*100, fill = NAME),
                  alpha = 0.2) +
      geom_line(aes(x = Year, y = Median*100),
                linewidth = 1.1,
                colour = "#4D4D4D") +
      geom_text(
        aes(x = Year,
            y = Upper*100 + 3,
            label = arrow_symbol),
        size = 6,
      family = "Arial Unicode MS"  # or "Segoe UI Symbol"
      ) +
      labs(
        title = .y,
        x = "Year",
        y = "Coral cover (%)"
      ) +
     # ylim(0, 70) +
      theme_pubr() +
      theme(
        axis.text  = element_text(size = 13),
        axis.title = element_text(size = 15),
        plot.title = element_text(size = 16),
        legend.position = "none"
      )   +
  scale_fill_manual("Island", values = c(
  "Tutuila"       = "#5b2a86",  
  "Ofu-Olosega"   = "#2c7fb8", 
  "Ta‘ū"          = "#41b6c4", 
  "Rose Atoll"    = "#7fcdbb",  
  "Swains Island" = "#f768a1"
   )) 
  })

p1 <- plots[[1]]
p2 <- plots[[2]]
p3 <- plots[[3]]
p4 <- plots[[4]]
p5 <- plots[[5]]

p_island <- p5 + p1 + p4 + p2 + p3 +
   plot_layout(nrow = 2, axis_titles = "collect") +  
   plot_annotation(tag_levels = "a", tag_suffix = ')', 
  title = ""
)

info_island <- pred_island$info_region

ggsave(p1, filename = "../figures/island_trend1.png",
        width=6, height=5)
ggsave(p2, filename = "../figures/island_trend2.png",
        width=6, height=5)
ggsave(p3, filename = "../figures/island_trend3.png",
        width=6, height=5) 
ggsave(p4, filename = "../figures/island_trend4.png",
        width=6, height=5) 
ggsave(p5, filename = "../figures/island_trend5.png",
        width=6, height=5) 
```

```{r}
#| include: false
#| echo: false
#| eval: false

#### Regional NRM scale {.unnumbered}

pred_NRM <- spatial_agg(post_dist_df_tier5,  post_dist_df_all, tiers.lookup, tier.sf, "Tier4")


plots <- pred_NRM$pred_table %>%
  droplevels() %>%
  mutate(
    Tier4_NRM = case_when(
      Tier4 == "1786" ~ "American Samoa",
      TRUE ~ NA_character_
    ),
    arrow_symbol = case_when(
      Change == "Up"   ~ "↑",
      Change == "Down" ~ "↓",
      Change == "Flat" ~ "→",
      TRUE ~ ""
    )
  ) %>%
  split(.$Tier4_NRM) %>%
  imap(~{
    ggplot(.x) +
      geom_ribbon(aes(x = Year, ymin = Lower*100, ymax = Upper*100),
                  alpha = 0.2,
                  fill = "#9B2226") +
      geom_line(aes(x = Year, y = Median*100),
                linewidth = 1.1,
                colour = "#4D4D4D") +
      geom_text(
        aes(x = Year,
            y = Upper*100 + 3,
            label = arrow_symbol),
        size = 6,
      family = "Arial Unicode MS"  # or "Segoe UI Symbol"
      ) +
      labs(
        title = .y,
        x = "Year",
        y = "Coral cover (%)"
      ) +
     # ylim(0, 70) +
      theme_pubr() +
      theme(
        axis.text  = element_text(size = 13),
        axis.title = element_text(size = 15),
        plot.title = element_text(size = 16),
        legend.position = "none"
      ) 
  })
p4 <- plots[[1]]

info_NRM <- pred_NRM$info_region

ggsave(p4, filename = "../figures/ecoregion_trend.png",
        width=6, height=5)

```

```{r, fig.width=8, fig.height=6}
#| include: true
#| echo: false
#| eval: false
#| label: fig-pred-island
#| fig-cap: Estimated coral trends at the island-scale from 2015 to 2023. Black lines represent mean predicted coral cover, with shaded areas indicating predictive intervals.
p_island
```


```{r, fig.width=8, fig.height=6}
#| include: true
#| echo: false
#| eval: false
#| label: fig-pred-region
#| fig-cap: Estimated coral trends at the ecoregion-scale from 2015 to 2023. Black lines represent mean predicted coral cover, with shaded areas indicating predictive intervals.
p4
```


```{r}
#| include: false
#| echo: false
#| eval: true

#### Overall attribution
coef_plots <- purrr::imap(mod_, function(mod, model_name) {
  get_coef_data(mod) %>%
    label_params() 
})

coef_all <- imap_dfr(coef_plots, ~ .x %>%
                       mutate(region = .y)) %>%
                       mutate(region_name = sub("_.*", "", region)) %>%
  mutate(neg_ci = if_else(effect_low < 0 & effect_high < 0, TRUE, FALSE)) 

# Reorder param by average effect_50 (most negative first)
param_order <- coef_all %>%
  group_by(param) %>%
  summarise(mean_effect = mean(effect_50, na.rm = TRUE)) %>%
  arrange(mean_effect) %>%
  pull(param)

coef_all <- coef_all %>%
  mutate(param = factor(param, levels = param_order))


# Prepare stripe data (one row per param x region_name)
stripe_data <- coef_all %>%
  distinct(param) %>%
  arrange(param) %>%
  mutate(y = row_number()/3)

g <- ggplot() +
  ggforestplot::geom_stripes(
    data = stripe_data,
    aes(y = param),
    odd = "white", even = "gray90"
  ) +
  geom_point(
    data = coef_all,
    aes(x = effect_50, y = param),
    size = 2.3, col = "#9B2226"
  ) +
  geom_errorbar(
    data = coef_all,
    aes(xmin = effect_low, xmax = effect_high, y = param),
    width = 0.1, col = "#9B2226"
  ) +
  geom_text(
    data = filter(coef_all, neg_ci),
    aes(x = effect_low - 0.005, y = param),
    label = "*", colour = "black", size = 4
  ) +
  theme_pubr() +
  geom_vline(xintercept = 0, linetype = "dashed", size = 0.4) +
  theme(legend.position = "bottom") +
  labs(x = "Effect Size", y = "")

g
 ggsave(g, filename = "../figures/Dist_effect_size.png",
        width=7, height=6)
```

```{r}
#| include: false
#| echo: false
#| eval: true

all_plots <- imap(mod_, ~plot_conditional_time(.x, .y))

plots <- all_plots %>%
  map(~ map(.x, "p")) %>%
  flatten()

declines_ASM <- map_dfr(all_plots[["FRK_shallow"]],
                    ~ compute_decline(.x$cond_table),
                    .id = "disturbance") %>%
  mutate(across(contains("decline"), ~ .x * 100))
all_plots_1 <- wrap_plots(plots, ncol = 2)

```

```{r}
#| include: false
#| echo: false
#| eval: true


# Conditional plots
# American Samoa

mod <- mod_[[1]]

mod_M <- mod$M
BAUs <- mod$obj_frk$ST_BAUs
obj_frk <- mod$obj_frk
hexpred <- mod$HexPred_reefid2
data_tier <- mod$data.grp.tier
sf::sf_use_s2(FALSE)

olddata <- mod_M@BAUs@data 
N <- nrow(olddata)
unique_hex_island <- plot_islands_custom_buffer(add_islands)$unique_hex_island

 newpred <- predict(mod_M, new_BAU_data = olddata)
  
  post_dist_df <- as.data.frame(newpred$MC$mu_samples) %>% 
    mutate(fYEAR = obj_frk$ST_BAUs@data$fYEAR) %>%
    mutate(Tier5 = obj_frk$ST_BAUs@data$Tier5) %>%
    mutate(id_loc = row_number()) %>%
    tidyr::pivot_longer(!c(fYEAR, Tier5, id_loc),
                        names_to = "draw", 
                        values_to = "pred")    
  
  # Spatial maps 
  hexpred$Tier5 <- as.factor(hexpred$Tier5)
  
  pred_sum_sf <- post_dist_df %>% 
    group_by(fYEAR, Tier5) %>% 
    median_hdci(pred) %>%
    inner_join(hexpred %>% 
                 group_by(Tier5) %>% 
                 summarize() %>% 
                 dplyr::select(geometry, Tier5)) %>% 
    st_as_sf(sf_column_name = "geometry") %>%
    mutate(Unc = .upper - .lower,
           tier_fYEAR = paste0(Tier5, fYEAR)) %>%
    dplyr::select(fYEAR, Tier5, pred, .lower, .upper, Unc, tier_fYEAR) %>%
  left_join(
    unique_hex_island %>% st_drop_geometry(),
    by = "Tier5"
  )
  
  # palettes
  pal_pred <- LaCroixColoR::lacroix_palette("Pamplemousse", n = 100, type = "continuous")
  pal_unc  <- wesanderson::wes_palette("Zissou1", 100, type = "continuous")
  
  # ---- split by island ----
  plots <- pred_sum_sf %>%
    split(.$NAME) %>%
    purrr::imap(function(df_island, island_name) {
      
      # prediction plot
      p <- ggplot() + 
        geom_sf(data = df_island, aes(fill = pred * 100), col = "transparent") +
        coord_sf(expand = FALSE) +
        scale_fill_gradientn(
          colours = pal_pred,
          name = "Coral Cover (%)",
          limits = c(0, 70)
        ) +
        facet_wrap(~ fYEAR) +
        theme_pubr() +
        labs(
          title = "",
          x = "Longitude",
          y = "Latitude"
        ) +
        theme(
          legend.position = "top",
          legend.title = element_text(size = 12),
          legend.text = element_text(size = 10),
          strip.background = element_blank(),
          strip.text = element_text(size = 12),
          plot.title = element_text(size = 13, hjust = 0.5),
          axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 9),
          axis.title = element_text(size = 12)
        )
      
      # uncertainty plot
      p_unc <- ggplot() + 
        geom_sf(data = df_island, aes(fill = Unc * 100), col = "transparent") +
        coord_sf(expand = FALSE) +
        scale_fill_gradientn(
          colours = pal_unc,
          name = "Uncertainty range (%)",
          limits = c(0, 70)
        ) +
        facet_wrap(~ fYEAR) +
        theme_pubr() +
        labs(
          title = "",
          x = "Longitude",
          y = "Latitude"
        ) +
        theme(
          legend.position = "top",
          legend.title = element_text(size = 12),
          legend.text = element_text(size = 10),
          strip.background = element_blank(),
          strip.text = element_text(size = 12),
          plot.title = element_text(size = 13, hjust = 0.5),
          axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 9),
          axis.title = element_text(size = 12) 
        ) 
      list(p = p, p_unc = p_unc)
    })

p_Tutuila <- plots$Tutuila$p + plots$Tutuila$p_unc  +
  plot_layout(nrow = 2, axis_titles = "collect") +  
  plot_annotation(
    tag_levels = "a", 
    tag_suffix = ')', 
    title = "",
    theme = theme(
      plot.title = element_text(size = 20)))

p_OO <- plots$`Ofu-Olosega`$p + plots$`Ofu-Olosega`$p_unc  +
  plot_layout(nrow = 2, axis_titles = "collect") +  
  plot_annotation(
    tag_levels = "a", 
    tag_suffix = ')', 
    title = "",
    theme = theme(
      plot.title = element_text(size = 20)))

p_RA <- plots$`Rose Atoll`$p + plots$`Rose Atoll`$p_unc  +
  plot_layout(nrow = 2, axis_titles = "collect") +  
  plot_annotation(
    tag_levels = "a", 
    tag_suffix = ')', 
    title = "",
    theme = theme(
      plot.title = element_text(size = 20)))

p_SI <- plots$`Swains Island`$p + plots$`Swains Island`$p_unc  +
  plot_layout(nrow = 2, axis_titles = "collect") +  
  plot_annotation(
    tag_levels = "a", 
    tag_suffix = ')', 
    title = "",
    theme = theme(
      plot.title = element_text(size = 20)))

p_T <- plots[[4]]$p + plots[[4]]$p_unc  +
  plot_layout(nrow = 2, axis_titles = "collect") +  
  plot_annotation(
    tag_levels = "a", 
    tag_suffix = ')', 
    title = "",
    theme = theme(
      plot.title = element_text(size = 20)))

## Plot for the paper 
plots_tul <- pred_sum_sf %>%
    filter(NAME == "Tutuila") 

p_tul <- ggplot() + 
        geom_sf(data = plots_tul, aes(fill = pred * 100), col = "transparent") +
        coord_sf(expand = FALSE) +
        scale_fill_gradientn(
          colours = pal_pred,
          name = "Coral Cover (%)",
          limits = c(0, 70)
        ) +
        facet_wrap(~ fYEAR) +
        theme_pubr() +
        labs(
          title = "",
          x = "Longitude",
          y = "Latitude"
        ) +
        theme(
          legend.position = "top",
          legend.title = element_text(size = 12),
          legend.text = element_text(size = 10),
          strip.background = element_blank(),
          strip.text = element_text(size = 12),
          plot.title = element_text(size = 13, hjust = 0.5),
          axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 9),
          axis.title = element_text(size = 12)
        )  +
  scale_y_continuous(
    breaks = c(-14.26, -14.32, -14.38),
    labels = c("14.26°S", "14.32°S", "14.38°S")
  ) +
  scale_x_continuous(
    breaks = c(-170.80, -170.70, -170.60),
    labels = c("170.80°W","170.70°W", "170.60°W"))
      
# uncertainty plot
p_unc_tul <- ggplot() + 
        geom_sf(data = plots_tul, aes(fill = Unc * 100), col = "transparent") +
        coord_sf(expand = FALSE) +
        scale_fill_gradientn(
          colours = pal_unc,
          name = "Uncertainty range (%)",
          limits = c(0, 70)
        ) +
        facet_wrap(~ fYEAR) +
        theme_pubr() +
        labs(
          title = "",
          x = "Longitude",
          y = "Latitude"
        ) +
        theme(
          legend.position = "top",
          legend.title = element_text(size = 12),
          legend.text = element_text(size = 10),
          strip.background = element_blank(),
          strip.text = element_text(size = 12),
          plot.title = element_text(size = 13, hjust = 0.5),
          axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
          axis.text.y = element_text(size = 9),
          axis.title = element_text(size = 12) 
        )  +
  scale_y_continuous(
    breaks = c(-14.26, -14.32, -14.38),
    labels = c("14.26°S", "14.32°S", "14.38°S")
  ) +
  scale_x_continuous(
    breaks = c(-170.80, -170.70, -170.60),
    labels = c("170.80°W","170.70°W", "170.60°W"))

p_Tutuila_paper <- p_tul + p_unc_tul  +
  plot_layout(nrow = 2, axis_titles = "collect") +  
  plot_annotation(
    tag_levels = "a", 
    tag_suffix = ')', 
    title = "Tutuila",
    theme = theme(
      plot.title = element_text(size = 20)))

ggsave(p_Tutuila_paper, filename = "../figures/Predictions_Tutuila.png",
         width=6, height=10)

################## Cyclone exposure
# ## Low 
new_BAU_data <- olddata %>%
  mutate(max_cyc      = rnorm(N, 0, 0.001),
         max_dhw      = rnorm(N, 0, 0.001)) 

p_cyc_low <- predict_newdata(mod_M, new_BAU_data, hexpred, obj_frk, title_n = "0 Hours", max_cover = 42)

ggsave(p_cyc_low[[1]], filename = "../figures/Low_Cyclone_ASM.png",
         width=7, height=6)


# High 
new_BAU_data <- olddata %>%
  mutate(max_cyc      = rnorm(N, 15, 0.001),
         max_dhw      = rnorm(N, 0, 0.001)) 

p_cyc_high <- predict_newdata(mod_M, new_BAU_data, hexpred, obj_frk, title_n = "15 Hours", max_cover = 42)

ggsave(p_cyc_high[[1]], filename = "../figures/High_Cyclone_ASM.png",
         width=7, height=6)


combined_plot_ASM <- p_cyc_low[[1]] + p_cyc_high[[1]] + guide_area() +
   plot_layout(
    nrow = 2,
    guides = "collect",
    axis_titles = "collect",
    heights = c(1, 0.15)  
  )  +  
   plot_annotation(tag_levels = "a", tag_suffix = ')', 
  title = "American Samoa"
)

### Compute differences 

# Wet Tropics 

pred_low <- p_cyc_low[[2]]
pred_high <- p_cyc_high[[2]]

p_diff_ASM <- plot_diff(pred_high, pred_low, title_n = "American Samoa")
ggsave(p_diff_ASM , filename = "../figures/ASM_diff.png",
         width=7, height=6)
```

### Model predictions {.unnumbered}

#### Tutuila

```{r, fig.width=8, fig.height=12}
#| include: true
#| echo: false
#| eval: true
#| label: fig-pred-1
#| fig-cap:  a) Predicted coral cover values across Tutuila island and b) associated uncertainty range for each year between 2015-2023. Values correspond to the posterior means.   
p_Tutuila
```


#### Ofu-Olosega

```{r, fig.width=12, fig.height=12}
#| include: true
#| echo: false
#| eval: true
#| label: fig-pred-2
#| fig-cap:  a) Predicted coral cover values across Ofu-Olosega island and b) associated uncertainty range for each year between 2015-2023. Values correspond to the posterior means.    
p_OO
```

#### Ta‘ū

```{r, fig.width=12, fig.height=12}
#| include: true
#| echo: false
#| eval: true
#| label: fig-pred-3
#| fig-cap:  a) Predicted coral cover values across Ta‘ū island and b) associated uncertainty range for each year between 2015-2023. Values correspond to the posterior means.    
p_T 
```

#### Rose Atoll

```{r, fig.width=12, fig.height=12}
#| include: true
#| echo: false
#| eval: true
#| label: fig-pred-4
#| fig-cap:  a) Predicted coral cover values across Rose Atoll and b) associated uncertainty range for each year between 2015-2023. Values correspond to the posterior means.   
p_RA  
```

#### Swains Island

```{r, fig.width=12, fig.height=12}
#| include: true
#| echo: false
#| eval: true
#| label: fig-pred-5
#| fig-cap:  a) Predicted coral cover values across Swains Island and b) associated uncertainty range for each year between 2015-2023. Values correspond to the posterior means. 
p_SI 
```

### Association with Disturbances {.unnumbered}

In American Samoa, cyclone exposure was significantly negatively associated with regional coral cover, with 95% prediction intervals not overlapping zero (@fig-cond-dist-1). The absolute decline between the minimum and maximum observed cyclone exposure, corresponds to a 16.1% reduction in coral cover (95% prediction interval: −27.8% to -1.2%, @fig-cond-all-1). 

Spatially, increasing magnitude of cyclone exposure alone (i.e., excluding heat stress) contributed to regional declines in coral cover (@fig-cond-space-1), with absolute losses estimated to range from 4.1% to 27.0%, and a regional average decline of 15.8% (@fig-cond-space-change-1).  

```{r, fig.width=5, fig.height=4}
#| include: true
#| echo: false
#| eval: true
#| label: fig-cond-dist-1
#| fig-cap: Predicted effect sizes of disturbances. The dots represent the predicted averaged effect and the intervals represent the corresponding 95\% predictive intervals. Stars show significant effect sizes when values of predictive intervals are entirely negatives. 
g
```


```{r, fig.width=10, fig.height=6}
#| include: true
#| echo: false
#| eval: true
#| label: fig-cond-all-1
#| fig-cap: Attribution of disturbances. The panels display the predicted coral cover across the range of disturbance values, from the minimum to the maximum values observed in American Samoa. Note that values of Heat Stress are expressed in Degree Heating Weeks and Cyclone Exposure in hours. The lines correspond to the averaged predicted values and shaded areas their associated 95 % prediction intervals. Effects have been calculated at the mean values of all the other covariates.   
all_plots_1
```

```{r, fig.width=10, fig.height=8}
#| include: true
#| echo: false
#| eval: true
#| label: fig-cond-space-1
#| fig-cap: Spatial predictions of coral cover under fixed low and high levels of cyclone exposure. 
combined_plot_ASM 
```

```{r, fig.width=8, fig.height=7}
#| include: true
#| echo: false
#| eval: true
#| label: fig-cond-space-change-1
#| fig-cap: Associated relative changes of coral cover between low and high levels of cyclone exposure.
p_diff_ASM
```

### Model validation {.unnumbered}

#### Model fit {.unnumbered}

```{r}
#| include: false
#| echo: false
#| eval: true

model_results <- imap(mod_, compute_model_fit)

plots <- purrr::map(model_results, "plot")

combined_plot_fit <- wrap_plots(plots, ncol = 1) +  
  plot_layout(guides = "collect", axis_titles = "collect") +  
  plot_annotation(tag_levels = "a", tag_suffix = ')')

ggsave(combined_plot_fit, filename = "../figures/models_fit.png",
        width=6, height=8)
```
```{r, fig.width=6, fig.height=6}
#| label: r-squared
#| eval: true
#| echo: false
#| fig-cap:  Model goodness-of-fit for the three regions. Coral cover observations at the tier-level were used to fit the spatio-temporal models. The R² statistic measures how well the predicted values reproduce the observations, with values closer to 1 indicating a better fit along the 1:1 red dotted line.    
combined_plot_fit
``` 

#### Model residuals {.unnumbered}

```{r}
#| include: false
#| echo: false
#| eval: true

res_list <- imap(mod_, ~ make_FRK_dharma_res(.x))
res1 <- gg_dharma(res_list[[1]], "") + ggtitle("") +
    theme(plot.title = element_text(size = 10))
ggsave(res1, filename = "../figures/model_res.png",
        width=8, height=6)
```

```{r, fig.width=8, fig.height=6}
#| label: mod-res
#| eval: true
#| echo: false
#| fig-cap: QQ plot residuals diagnostic.     
res1
``` 
